---
title: "Take-home Exercise 1"
author: "XieHaoyang"
date-modified: "last-modified"
execute: 
  warning: false
---

# 1. Overview

## 1.1 Background

## 1.2 Objectives

# 2. Data Wrangling

## 2.1 Loading Packages

```{r}
pacman::p_load(sf, tmap, tidyverse, httr, ggplot2, dplyr, lubridate, spatstat, stars)
```

## 2.2 The Data

The datasets are Information on Corporate Entities datasets of ACRA. There are a total of 27 csv files organised alphabetically.

To facilitate data extraction, the original datasets have been merged to generate a new dataset using the following code.

```         
folder_path <- "data/aspatial/acra"
file_list <- list.files(path = folder_path,
                        pattern = "^ACRA*.*\\.csv$",
                        full.names = TRUE)

acra_data <- file_list %>%
  map_dfr(read.csv)

write_rds(acra_data,
          "data/rds/acra_data.rds")
```

```         
acra_data <- read_rds("D:/XieHaoyang3537/ISSS626-GAA-XieHaoyang/In-class_Ex/In-class_ex2/data/rds/acra_data.rds")
```

## 2.3 Geocoding

Time Range

```         
start_date <- as.Date("2024-01-01")
end_date   <- as.Date("2025-06-30")
```

Define functions: filtering + cleaning + geocoding

```         
prep_biz_data <- function(df, ssic_codes) {
  
  biz <- df %>%
    select(1:24) %>%
    filter(primary_ssic_code %in% ssic_codes) %>%
    rename(date = registration_incorporation_date) %>%
    mutate(date = as.Date(date)) %>%
    filter(date >= start_date & date <= end_date) %>%
    mutate(
      YEAR = year(date),
      MONTH_NUM = month(date),
      MONTH_ABBR = month(date, label = TRUE, abbr = TRUE),
      postal_code = str_pad(postal_code, width = 6, side = "left", pad = "0")
    )
  
  postcodes <- unique(biz$postal_code)
  url <- "https://onemap.gov.sg/api/common/elastic/search"
  
  found <- data.frame()
  not_found <- data.frame(postcode = character())
  
  for (pc in postcodes) {
    query <- list(
      searchVal = pc,
      returnGeom = "Y",
      getAddrDetails = "Y",
      pageNum = "1"
    )
    
    res <- GET(url, query = query)
    json <- content(res)
    
    if (json$found != 0) {
      df_res <- as.data.frame(json$results, stringsAsFactors = FALSE)
      df_res$input_postcode <- pc
      found <- bind_rows(found, df_res)
    } else {
      not_found <- bind_rows(not_found, data.frame(postcode = pc))
    }
  }
  
  biz <- biz %>%
    left_join(found, by = c("postal_code" = "POSTAL"))
  
  biz_sf <- st_as_sf(biz, coords = c("X", "Y"), crs = 3414)
  
  return(biz_sf)
}
```

Select the industry and save the file

```         
ssic_fnb   <- c(56111, 56112, 56113)  # Cafes, Food courts, Restaurants
ssic_ict   <- c(62021, 63120, 58201)  # IT consultancy, web portals, software publishing
ssic_edu   <- c(85410, 85509)         # Tuition, other education

biz_fnb <- prep_biz_data(acra_data, ssic_fnb)
biz_ict <- prep_biz_data(acra_data, ssic_ict)
biz_edu <- prep_biz_data(acra_data, ssic_edu)

write_rds(biz_fnb, "data/rds/biz_fnb.rds")
write_rds(biz_ict, "data/rds/biz_ict.rds")
write_rds(biz_edu, "data/rds/biz_edu.rds")
```

# 3. Mapping the geospatial data sets

Importing data

```{r}
biz_fnb <- read_rds("data/rds/biz_fnb.rds")
biz_ict <- read_rds("data/rds/biz_ict.rds")
biz_edu <- read_rds("data/rds/biz_edu.rds")

sg_boundary <- st_read("data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>% 
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```

## 3.1 F&B Businesses

```{r}
tmap_mode("plot")
tm_shape(sg_boundary) +
  tm_polygons(col = "lightgrey", border.col = "white") +
  tm_shape(biz_fnb) +
  tm_dots(col = "blue", size = 0.2, alpha = 0.6) +
  tm_layout(title = "F&B Businesses (2024-2025)")
```

## 3.2 Information and Communications Industry

```{r}
tmap_mode("plot")
tm_shape(sg_boundary) +
  tm_polygons(col = "lightgrey", border.col = "white") +
  tm_shape(biz_ict) +
  tm_dots(col = "red", size = 0.2, alpha = 0.6) +
  tm_layout(title = "Information and Communications Industry (2024-2025)")
```

## 3.3 Education

```{r}
tmap_mode("plot")
tm_shape(sg_boundary) +
  tm_polygons(col = "lightgrey", border.col = "white") +
  tm_shape(biz_edu) +
  tm_dots(col = "green", size = 0.2, alpha = 0.6) +
  tm_layout(title = "Education (2024-2025)")
```

### KDE

```{r}
biz_fnb_ppp <- as.ppp(st_coordinates(biz_fnb), 
                      W = as.owin(st_union(sg_boundary)))

kde_fnb <- density(biz_fnb_ppp, sigma = 1000)

kde_fnb_stars <- st_as_stars(kde_fnb)
st_crs(kde_fnb_stars) <- 3414  

tm_shape(kde_fnb_stars) + 
  tm_raster(alpha = 0.7, palette = "Blues") +
  tm_shape(sg_boundary) + tm_borders() +
  tm_layout(title = "KDE Hotspots - F&B Businesses (2024-2025)")
```

```{r}
biz_ict_ppp <- as.ppp(st_coordinates(biz_ict), 
                      W = as.owin(st_union(sg_boundary)))

kde_ict <- density(biz_ict_ppp, sigma = 1000)

kde_ict_stars <- st_as_stars(kde_ict)
st_crs(kde_ict_stars) <- 3414  

tm_shape(kde_ict_stars) + 
  tm_raster(alpha = 0.7, palette = "Reds") +
  tm_shape(sg_boundary) + tm_borders() +
  tm_layout(title = "KDE Hotspots - Information and Communications Industry (2024-2025)")
```

```{r}
biz_edu_ppp <- as.ppp(st_coordinates(biz_edu), 
                      W = as.owin(st_union(sg_boundary)))

kde_edu <- density(biz_edu_ppp, sigma = 1000)

kde_edu_stars <- st_as_stars(kde_edu)
st_crs(kde_edu_stars) <- 3414  

tm_shape(kde_edu_stars) + 
  tm_raster(alpha = 0.7, palette = "Greens") +
  tm_shape(sg_boundary) + tm_borders() +
  tm_layout(title = "KDE Hotspots - Education (2024-2025)")
```

## Distribution in the Time Dimension
