---
title: "Hands-on exercise 2-2: 2nd Order Spatial Point Patterns Analysis Methods"
author: "XieHaoyang"
---

## Getting Started

### Install and lauching R packages

```{r}
pacman::p_load(sf, spatstat, tmap, tidyverse, rvest)
```

## Data Import and Preparation

```{r}
childcare_tm_ppp <- readRDS("data/childcare_tm_ppp.rds")
childcare_ck_ppp <- readRDS("data/childcare_ck_ppp.rds")
```

## Second-order Spatial Point Patterns Analysis

## Analysing Spatial Point Process Using G-Function

### Choa Chu Kang planning area

#### Computing G-function estimation

```{r}
set.seed(1234)
```

```{r}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```

#### Performing Complete Spatial Randomness Test

```{r}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)
```

```{r}
plot(G_CK.csr)
```

### Tampines planning area

#### Computing G-function estimation

```{r}
G_tm = Gest(childcare_tm_ppp, correction = "best")
plot(G_tm)
```

#### Performing Complete Spatial Randomness Test

```{r}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)
```

```{r}
plot(G_tm.csr)
```

## Analysing Spatial Point Process Using F-Function

### Choa Chu Kang planning area

#### Computing F-function estimation

```{r}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```

### Performing Complete Spatial Randomness Test

```{r}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)
```

```{r}
plot(F_CK.csr)
```

### Tampines planning area

#### Computing F-function estimation

```{r}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```

#### Performing Complete Spatial Randomness Test

```{r}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)
```

```{r}
plot(F_tm.csr)
```

## Analysing Spatial Point Process Using K-Function

### Choa Chu Kang planning area

#### Computing K-fucntion estimate

```{r}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

#### Performing Complete Spatial Randomness Test

```{r}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

### Tampines planning area

#### Computing K-fucntion estimation

```{r}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

#### Performing Complete Spatial Randomness Test

```{r}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))
```

## Analysing Spatial Point Process Using L-Function

### Choa Chu Kang planning area

#### Computing L Fucntion estimation

```{r}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

#### Performing Complete Spatial Randomness Test

```{r}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

### Tampines planning area

#### Computing L-fucntion estimate

```{r}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

#### Performing Complete Spatial Randomness Test

```{r}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```
