---
title: "Hands-on exercise 5: Global Measures of Spatial Autocorrelation"
author: "XieHaoyang"
---

## Getting Started

### Setting the Analytical Tools

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse)
```

## Getting the Data Into R Environment

### Import shapefile into r environment

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### Import csv file into r environment

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### Performing relational join

```{r}
hunan <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)
```

### Visualising Regional Development Indicator

```{r}
equal <- tm_shape(hunan) +
  tm_polygons(fill = "GDPPC",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 5,
                values = "brewer.blues")) +
  tm_borders(fiil_alpha = 0.5) + 
  tm_layout(legend.position = c("left", "bottom"), 
            main.title = "Equal interval classification")

quantile <- tm_shape(hunan) +
  tm_polygons(fill = "GDPPC",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5)) +
  tm_borders(fiil_alpha = 0.5) + 
  tm_layout(legend.position = c("left", "bottom"), 
            main.title = "Quantile classification")

tmap_arrange(equal, 
             quantile, 
             asp=1, 
             ncol=2)
```

## Global Measures of Spatial Autocorrelation

### Computing Contiguity Spatial Weights

```{r}
wm_q <- poly2nb(hunan, 
                queen=TRUE)
summary(wm_q)
```

### Row-standardised weights matrix

```{r}
rswm_q <- nb2listw(wm_q, 
                   style="W", 
                   zero.policy = TRUE)
rswm_q
```

## Global Measures of Spatial Autocorrelation: Moran’s I

### Maron’s I test

```{r}
moran.test(hunan$GDPPC, 
           listw=rswm_q, 
           zero.policy = TRUE, 
           na.action=na.omit)
```

### Computing Monte Carlo Moran’s I

```{r}
set.seed(1234)
bperm= moran.mc(hunan$GDPPC, 
                listw=rswm_q, 
                nsim=999, 
                zero.policy = TRUE, 
                na.action=na.omit)
bperm
```

### Visualising Monte Carlo Moran’s I

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, 
     freq=TRUE, 
     breaks=20, 
     xlab="Simulated Moran's I")
abline(v=0, 
       col="red") 
```

## Global Measures of Spatial Autocorrelation: Geary’s C

### Geary’s C test

```{r}
geary.test(hunan$GDPPC, listw=rswm_q)
```

### Computing Monte Carlo Geary’s C

```{r}
set.seed(1234)
bperm=geary.mc(hunan$GDPPC, 
               listw=rswm_q, 
               nsim=999)
bperm
```

### Visualising the Monte Carlo Geary’s C

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, freq=TRUE, breaks=20, xlab="Simulated Geary c")
abline(v=1, col="red") 
```

## Spatial Correlogram

### Compute Moran’s I correlogram

```{r}
MI_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="I", 
                          style="W")
plot(MI_corr)
```

```{r}
print(MI_corr)
```

### Compute Geary’s C correlogram and plot

```{r}
GC_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="C", 
                          style="W")
plot(GC_corr)
```

```{r}
print(GC_corr)
```
