---
title: "Take-home Exercise 3"
author: "XieHaoyang"
date-modified: "last-modified"
execute: 
  warning: false
---

# 1. Overview

## 1.1 Background

## 1.2 Objectives

# 2. Data Wrangling

## 2.1 Loading Packages

```{r}
pacman::p_load(httr, tidyverse, sf, tmap, jsonlite, progress, rsample, GWmodel)
```

## 2.2 The Data

## 2.3 Importing HDB data

```{r,eval=FALSE}
HDBresale_raw <- read_csv("data/rawdata/HDBresale.csv")
```

```{r,eval=FALSE}
HDBresale <- HDBresale_raw %>%
  filter(flat_type == "4 ROOM") %>%
  filter(month >= "2025-01" & month < "2025-10")

head(HDBresale)
```

## 2.4 Geocoding

### 2.4.1 Preprocess street name

```{r,eval=FALSE}
HDBresale <- HDBresale %>%
  mutate(street_name = gsub("ST\\.", "SAINT", street_name))
```

### 2.4.2 Define the geocode function

```{r,eval=FALSE}
geocode <- function(block, streetname) {
  base_url <- "https://onemap.gov.sg/api/common/elastic/search"
  address <- paste(block, streetname, sep = "")
  query <- list(
    "searchVal" = address,
    "returnGeom" = "Y",
    "getAddrDetails" = "N",
    "pageNum" = "1"
  )
  
  res <- GET(base_url, query = query)
  restext <- content(res, as = "text")
  json <- fromJSON(restext)
  
  if(length(json$results) == 0){
    return(data.frame(LATITUDE = NA, LONGITUDE = NA))
  }
  
  output <- as.data.frame(json$results) %>%
    select(LATITUDE, LONGITUDE)
  
  return(output[1, ])
}
```

### 2.4.3 Data geocoding

```{r,eval=FALSE}
HDBresale$LATITUDE <- NA
HDBresale$LONGITUDE <- NA

for(i in 1:nrow(HDBresale)){
  temp_output <- tryCatch({
    geocode(HDBresale[i,"block"], HDBresale[i,"street_name"])
  }, error = function(e){
    data.frame(LATITUDE = NA, LONGITUDE = NA)
  })
  
  HDBresale$LATITUDE[i] <- temp_output$LATITUDE
  HDBresale$LONGITUDE[i] <- temp_output$LONGITUDE
}
```

```{r,eval=FALSE}
write_rds(HDBresale, "data/rds/HDB.rds")
```

```{r}
HDB <- readRDS("data/rds/HDB.rds")
```

### 2.4.4 Convert SF data

```{r}
HDBresale_sf <- HDB %>%
  st_as_sf(
    coords = c("LONGITUDE", "LATITUDE"),
    crs = 4326) %>%
  st_transform(crs = 3414)
```

## 2.5 Load other geospatial data

```{r}
mpsz <- st_read("data/rawdata/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>% 
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)

preschool = st_read("data/rawdata/PreschoolsLocation.kml") %>%
  st_transform(crs = 3414)

mrt <- st_read("data/rawdata/LTAMRTStationExitKML.kml") %>% 
  st_transform(crs = 3414)
```

## 2.6 Distance calculation

```{r}
# Calculate the distance from each HDB unit to the nearest MRT station.
nearest_mrt <- st_nearest_feature(HDBresale_sf, mrt)
HDBresale_sf$PROX_MRT <- st_distance(HDBresale_sf, mrt[nearest_mrt, ], by_element = TRUE)
# Convert to kilometers
HDBresale_sf$PROX_MRT <- as.numeric(HDBresale_sf$PROX_MRT) / 1000

# Calculate the number of preschools within 350 meters of each HDB unit.
within_350m_preschool <- st_is_within_distance(HDBresale_sf, preschool, dist = 350)
HDBresale_sf$WITHIN_350M_KINDERGARTEN <- map_int(within_350m_preschool, length)

head(HDBresale_sf %>% select(PROX_MRT, WITHIN_350M_KINDERGARTEN))
```

## 3. Exploratory analysis

### 3.1 Spatial distribution of HDB resale prices

```{r}
tmap_mode("plot")

tm_shape(mpsz) +
  tm_polygons(alpha = 0.1, 
              border.col = "grey") +
tm_shape(HDBresale_sf) +
  tm_dots(col = "resale_price",
          palette = "viridis",
          style = "quantile",
          n = 5,
          size = 0.2,
          alpha = 0.8,
          title = "HDB Resale Price (SGD)") +
  tm_layout(title = "Spatial Distribution of 4-Room HDB Resale Prices (2025)",
            frame = FALSE) 
```

### 3.2 Spatial pattern visualizing distance to MRT

```{r}
tm_shape(mpsz) +
  tm_polygons(alpha = 0.1,
              border.col = "grey") +
tm_shape(HDBresale_sf) +
  tm_dots(col = "PROX_MRT",
          palette = "-plasma",
          style = "quantile",
          n = 5,
          size = 0.2,
          alpha = 0.8,
          title = "Distance to Nearest MRT (km)") +
tm_shape(mrt) + 
  tm_dots(col = "red",
          size = 0.2,
          alpha = 0.8,
          title = "MRT Station") +
  tm_layout(title = "Spatial Distribution: Distance to MRT Stations",
            frame = FALSE)
```

### 3.3 Spatial pattern of preschool density visualization

```{r}
tm_shape(mpsz) +
  tm_polygons(alpha = 0.1,
              border.col = "grey") +
tm_shape(HDBresale_sf) +
  tm_dots(col = "WITHIN_350M_KINDERGARTEN",
          palette = "Blues",
          style = "fixed",
          breaks = c(0, 3, 6, 9, 12, 15),
          size = 0.2,
          alpha = 0.8,
          title = "Number of Preschools within 350m") +
tm_shape(preschool) + 
  tm_dots(col = "purple",
          size = 0.2,
          alpha = 0.6,
          title = "Preschool") +
  tm_layout(title = "Spatial Distribution: Preschool Density",
            frame = FALSE)
```

# 4. Modeling

## 4.1 Building a modeling dataset

### 4.1.1 Structural Factors

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    
    remaining_lease_mths = ifelse(
      "remaining_lease" %in% names(.),
      as.numeric(remaining_lease) * 12,  
      (99 - (2025 - as.numeric(substr(lease_commence_date, 1, 4)))) * 12  
    ),
    
    storey_order = case_when(
      storey_range == "01 TO 03" ~ 1,
      storey_range == "04 TO 06" ~ 2,
      storey_range == "07 TO 09" ~ 3,
      storey_range == "10 TO 12" ~ 4,
      storey_range == "13 TO 15" ~ 5,
      storey_range == "16 TO 18" ~ 6,
      storey_range == "19 TO 21" ~ 7,
      storey_range == "22 TO 24" ~ 8,
      storey_range == "25 TO 27" ~ 9,
      storey_range == "28 TO 30" ~ 10,
      storey_range == "31 TO 33" ~ 11,
      storey_range == "34 TO 36" ~ 12,
      storey_range == "37 TO 39" ~ 13,
      storey_range == "40 TO 42" ~ 14,
      TRUE ~ 7  
    )
  )
```

### 4.1.2 Location Factors

```{r}
# Calculate the distance to the CBD (assuming the CBD is near Raffles Place).
cbd_point <- data.frame(lon = 103.851959, lat = 1.284053) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(crs = 3414)

HDBresale_sf$PROX_CBD <- st_distance(HDBresale_sf, cbd_point) %>% 
  as.numeric() / 1000  
```

```{r}
within_1km_pri <- st_is_within_distance(HDBresale_sf, preschool, dist = 1000)
HDBresale_sf$WITHIN_1KM_PRISCH <- map_int(within_1km_pri, length)
```

## 4.1.3 Data quality check

```{r}
# Check for missing values
cat("=== Data integrity check ===\n")
cat("Total number of records:", nrow(HDBresale_sf), "\n\n")
cat("Missing value statistics:\n")
missing_summary <- sapply(HDBresale_sf %>% st_drop_geometry(), function(x) sum(is.na(x)))
print(missing_summary[missing_summary > 0])  

# Examine summary statistics of key variables
cat("\n=== Summary statistics of key variables ===\n")
HDBresale_sf %>% 
  st_drop_geometry() %>%
  select(resale_price, floor_area_sqm, remaining_lease_mths, storey_order, 
         PROX_MRT, PROX_CBD, WITHIN_350M_KINDERGARTEN, WITHIN_1KM_PRISCH) %>%
  summary()
```
The data has no missing values, making it suitable for modeling. However, we noticed that the remaining lease term value was not identified, so further processing is needed.

```{r}
# Convert "XX years YY months" to month values.
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    # Extract year numbers
    lease_years = as.numeric(str_extract(remaining_lease, "\\d+(?=\\s*years)")),
    # Extract month numbers
    lease_months = as.numeric(str_extract(remaining_lease, "\\d+(?=\\s*months)")) %>% replace_na(0),
    # Calculate the total number of months
    remaining_lease_mths = lease_years * 12 + lease_months
  ) %>%
  # Remove temporary variables
  select(-lease_years, -lease_months)

# Verify conversion results
cat("\nTransformed remaining_lease_mths summary:\n")
summary(HDBresale_sf$remaining_lease_mths)

# Recheck missing values
cat("\n=== Final missing value check ===\n")
sapply(HDBresale_sf %>% st_drop_geometry(), function(x) sum(is.na(x)))
```

## 4.2 Data sampling and segmentation

```{r}
set.seed(1234)

if(nrow(HDBresale_sf) > 1500) {
  HDB_sample <- HDBresale_sf %>%
    sample_n(1500)
} else {
  HDB_sample <- HDBresale_sf
}

cat("Data volume after sampling:", nrow(HDB_sample), "records\n")

# Check for overlapping points in the sampled data.
overlapping_points <- HDB_sample %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
cat("Number of overlapping points:", sum(overlapping_points$overlap), "\n")

# Split the training set and the test set
set.seed(1234)
resale_split <- initial_split(HDB_sample, prop = 0.67)
train_data <- training(resale_split)
test_data <- testing(resale_split)

cat("Training set data volume:", nrow(train_data), "records\n")
cat("Test set data size:", nrow(test_data), "records\n")
```

## 4.2.1 Handling overlapping points

```{r}
# Add random offsets to overlapping points
set.seed(1234)
train_data <- train_data %>%
  group_by(geometry) %>%
  mutate(
    n_duplicates = n(),

    geometry = ifelse(n_duplicates > 1,
                     geometry + runif(2, -1, 1),
                     geometry)
  ) %>%
  ungroup() %>%
  st_sf()

cat("After handling overlapping points, the number of unique locations in the training set:", nrow(unique(train_data)), "\n")
```

## 4.3 OLS model

```{r}
# Calibrate OLS model
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_MRT + 
                  WITHIN_350M_KINDERGARTEN + WITHIN_1KM_PRISCH,
                data = train_data)

# Display OLS results
cat("=== OLS model results ===\n")
summary(price_mlr)

# Check the model fit
cat("\n=== Model fit ===\n")
cat("R-squared:", summary(price_mlr)$r.squared, "\n")
cat("Adjusted R-squared:", summary(price_mlr)$adj.r.squared, "\n")
```

## 4.4 GWR model

### 4.4.1 Calibrate GWR model

```{r}
# Convert the sf object to a SpatialPointsDataFrame
train_data_sp <- as(train_data, "Spatial")

# Calculate the optimal bandwidth for GWR
cat("Start calculating GWR bandwidth\n")
bw_adaptive <- bw.gwr(
  resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths +
    PROX_CBD + PROX_MRT + 
    WITHIN_1KM_PRISCH,  
  data = train_data_sp,
  approach = "CV",
  kernel = "gaussian",
  adaptive = TRUE,
  longlat = FALSE
)

cat("Optimal bandwidth (number of neighboring points):", bw_adaptive, "\n")

# Save bandwidth results
write_rds(bw_adaptive, "data/bw_adaptive.rds")
```

### 4.4.2 Running the GWR model

```{r}
gwr_adaptive <- gwr.basic(
  resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths +
    PROX_CBD + PROX_MRT + 
    WITHIN_1KM_PRISCH,
  data = train_data_sp,
  bw = bw_adaptive,
  kernel = "gaussian",
  adaptive = TRUE,
  longlat = FALSE
)

cat("=== GWR model results ===\n")
print(gwr_adaptive)
```

```{r}
# Comparing the performance of OLS and GWR
cat("\n=== Model Comparison ===\n")
cat("OLS R-squared:", round(summary(price_mlr)$r.squared, 4), "\n")
cat("GWR R-squared:", round(gwr_adaptive$GW.diagnostic$gw.R2, 4), "\n")
cat("GWR Adjusted R-squared:", round(gwr_adaptive$GW.diagnostic$gwR2.adj, 4), "\n")
cat("GWR RSS:", round(gwr_adaptive$GW.diagnostic$RSS, 2), "\n")
cat("OLS RSS:", round(sum(price_mlr$residuals^2), 2), "\n")
```

```{r}
# Add the GWR results back to the training data.
train_data_with_gwr <- train_data %>%
  mutate(
    gwr_residuals = gwr_adaptive$SDF$residual,
    gwr_fitted = gwr_adaptive$SDF$yhat
  )

# Save GWR model results
write_rds(gwr_adaptive, "data/gwr_model.rds")
cat("The nGWR model has been saved to: data/gwr_model.rds\n")
```

## 4.5 Visualization of modeling results

```{r}
# Add the GWR coefficient to the spatial data.
gwr_results <- st_as_sf(gwr_adaptive$SDF) %>%
  st_set_crs(3414)

# Visualize the coefficients of key variables.
coefficients_to_plot <- c("floor_area_sqm", "storey_order", "PROX_CBD", "PROX_MRT")
```

```{r}
## Spatial distribution map of GWR coefficient
tmap_mode("plot")

for(coef in coefficients_to_plot) {
  map_title <- paste("GWR Coefficient:", coef)
  
  p <- tm_shape(mpsz) +
    tm_polygons(alpha = 0.1, border.col = "grey") +
    tm_shape(gwr_results) +
    tm_dots(col = coef,
            palette = "RdYlBu",
            style = "pretty",
            size = 0.2,
            title = paste("Coefficient:", coef),
            midpoint = 0) + 
    tm_layout(title = map_title,
              frame = FALSE)
  
  print(p)
}
```

## 4.6 Residual comparison

### 4.6.1 GWR residual space distribution

```{r}
train_data_with_gwr <- train_data %>%
  mutate(
    gwr_residuals = gwr_adaptive$SDF$residual,
    gwr_fitted = gwr_adaptive$SDF$yhat,
    ols_residuals = price_mlr$residuals
  )
```

```{r}
tm_shape(mpsz) +
  tm_polygons(alpha = 0.1, border.col = "grey") +
  tm_shape(train_data_with_gwr) +
  tm_dots(col = "gwr_residuals",
          palette = "RdYlBu",
          style = "pretty",
          size = 0.5,
          title = "GWR Residuals (SGD)",
          midpoint = 0) +
  tm_layout(title = "Spatial Distribution of GWR Residuals",
            frame = FALSE)
```

```{r}
# Save the final analysis dataset
write_rds(train_data_with_gwr, "data/final_analysis_data.rds")

# Final Model Summary
cat("\n=== Final Model Performance Summary ===\n")
cat("OLS R-squared:", round(summary(price_mlr)$r.squared, 4), "\n")
cat("GWR R-squared:", round(gwr_adaptive$GW.diagnostic$gw.R2, 4), "\n")
cat("Performance improvement:", round((gwr_adaptive$GW.diagnostic$gw.R2 - summary(price_mlr)$r.squared) * 100, 1), "%\n")
cat("Number of valid parameters for GWR:", round(gwr_adaptive$GW.diagnostic$enp, 1), "\n")
```
